---
title: "Downloading Retrosheet Data"
author: "Jim Albert"
date: "January 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here's an outline of how to download the Retrosheet play files for a particular season, and also how to compute the runs values for all plays.

There is a blog posting on this at
https://baseballwithr.wordpress.com/2014/02/10/downloading-retrosheet-data-and-runs-expectancy/ 

1.  Download the Chadwick files following the advice from the website mentioned on the blog post -- this will parse the original source files.

2.  Set up the current working directory to have a folder called "download.folder".  Inside that folder, place two subfolders "unzipped" and "zipped".  The following system commands will accomplish this.

```{r}
system("mkdir download.folder")
setwd("download.folder")
system("mkdir unzipped")
system("mkdir zipped")
setwd("..")
```

3.  Load the PitchSequence package.

```{r}
library(PitchSequence)
```

4.  Use the ```parse.retrosheet2.pbp()``` function to download the Retrosheet play-by-play data for the 2020 season.

```{r}
parse.retrosheet2.pbp(2020)
```

Navigate to the download file and check that three files are there. 

```{r}
setwd("download.folder/unzipped")
dir()
```

5.  To compute the runs expectancies, run the second function ```compute.runs.expectancy()```, saving the result in d2020.

```{r}
d2020 <- compute.runs.expectancy(2020)
```

6.  Check that you are getting sensible values:

Display the starting state, the new state and the runs value for the first few plays:

```
library(tidyverse)
d2020 %>% 
  select(STATE, NEW.STATE, RUNS.VALUE) %>% 
  head()
```

Display the runs expectancies for all states:

```
d2020 %>% group_by(STATE) %>% 
  summarize(R = first(RUNS.STATE))
```

Show the runs values for all possible transitions between states:

```
d2020 %>% group_by(STATE, NEW.STATE) %>% 
  summarize(R = first(RUNS.VALUE))
```



